---
title: 'Reproducible Research: Peer Assessment 1'
author: "VadimEv"
date: "16 may 2015"
output: html_document
---

### Loading and preprocessing data
In this chunk we load and preprocess data for analysis.
First, read the data from csv

```{r}
DT <-read.csv("activity.csv")
```
Now lets have a quick look at what we just loaded:
```{r}
head(DT)
```
So we see records of number of steps, date at which this record was recorded and interval, during which data was gathered.

### What is mean total number of steps taken per day?
In first part we calculate a number of steps taken each day
```{r}
DTdailysum <- aggregate(list(steps = DT$steps), list(date = DT$date), sum)
```
We can have a quick look at it, but not to make big textwalls present I use head, if you willing to, just replace head with view:
```{r}
head(DTdailysum)
```
lets calculate mean value of steps taken daily:
```{r}
mean(DTdailysum$steps, na.rm = TRUE)
```
and now we gonna make histogram plot of steps taken each day 

```{r}
hist(DTdailysum$steps, xlab ="total daily steps", main = "Total steps distribution")
```
</br>
As we are allready calculated mean, now only median value is needed:
```{r}
median(DTdailysum$steps, na.rm=TRUE)
```
### What is the average daily activity pattern?
To figure this out we gonna calculate number of steps in each 5 min time interval, averaged by day
```{r}
fsteps <-aggregate(data=DT, steps~interval, FUN = mean, na.action = na.omit)
```
and plot it with ggplot:
```{r}
library(ggplot2)
ggplot(data = fsteps, aes(interval, steps )) + geom_line()
```
</br>
now lets find the interval during which maximum number of steps was taken, first we gonna find value itself:
```{r}
maxstp <- max(fsteps$steps, na.rm=TRUE)
maxstp
```
and find which interval was it:
```{r}
fsteps[fsteps$steps == maxstp,]
```
so the interval is between 835th and 840 minute.

### Imputing missing values
Now we need to handle missing values, first of all we should count them:
```{r}
sum(is.na(DT$steps))
```
In my opinion, the best strategy is to replace all NA with zeros. Why should we invent something? 
```{r}
DT[is.na(DT)] <- 0
```
Lets have a look at how data changed with this, and calculate total number of steps taken each day, as well as median and mean values:
```{r}
DTdailysumN <- aggregate(list(steps = DT$steps), list(date = DT$date), sum)
mean(DTdailysumN$steps, na.rm = TRUE)
median(DTdailysumN$steps, na.rm = TRUE)
```
and make a histogram of this new data:
```{r}
hist(DTdailysumN$steps, xlab ="total daily steps", main = "Total steps distribution with NA removed")
```
</br>
As we can see, distribution changed according to addition of zero values. 

### Are there differences in activity patterns between weekdays and weekends?
To answer this question we need to introduce new column to the dataset, indicating if the day is weekend of not. Add a column for weekday:
```{r}
DT$weekday = weekdays(as.Date(DT$date))
```
and another column for weekday type, and convert it into factor:
```{r}
DT$type <- ifelse(DT$weekday == "Saturday" | DT$weekday == "Sunday", "Weekend", "Weekday")
DT$type <- as.factor(DT$type)
```
To see the differences lets compare daily activity on weekend and on a weekday, but before we need to aggregate data as before taking day type into account:
```{r}
fsteps <-aggregate(data=DT, steps~ type + interval, FUN = mean, na.action = na.omit)
ggplot(data = fsteps, aes(interval, steps )) + geom_line() + facet_grid(.~type)
```
</br>
GGplot2 works really well here, and we can clearly see that patterns differ pretty significantly.
